%---------------------------------------------------------------
%  SECTION V — CONCLUSION AND FUTURE WORK
%---------------------------------------------------------------
\section{Conclusion and Future Work}
\label{sec:conclusion}

This paper presented \textbf{Q-SFTP}, a production-grade post-quantum
secure file transfer protocol that replaces all classical cryptographic
primitives with NIST-standardised algorithms. The key contributions are:

\begin{enumerate}
  \item A four-phase Q-TLS handshake using Kyber512 (IND-CCA2) and
        Dilithium2 (EUF-CMA) achieving mutual authentication and
        ephemeral forward secrecy in under 80~ms.
  \item A BLAKE2b-256 dual-side integrity ecosystem with persistent
        hash registry and periodic background monitoring for at-rest
        tamper detection.
  \item A chunked resumable transfer protocol with $\approx 0.01\%$
        cryptographic overhead per file, surviving connection
        interruptions and session restarts.
  \item Defense-in-depth: three-tier RBAC with directory isolation,
        automatic metadata stripping, magic-number file validation,
        and tamper-evident audit logging.
  \item A browser-based web GUI and Docker containerisation enabling
        zero-configuration deployment.
\end{enumerate}

Experimental evaluation confirms quantum-safe security against all
considered attack classes (quantum cryptanalysis, MITM, replay, and
data tampering) with negligible performance overhead, demonstrating
that the migration from classical to post-quantum secure communication
is both achievable and practical today.

\subsection*{Future Work}

\begin{itemize}
  \item \textbf{Hybrid KEM}: Combining Kyber512 with X25519 for
        defence-in-depth against unforeseen lattice weaknesses
        \cite{chen2022hybridtls}.
  \item \textbf{Formal verification}: Applying ProVerif or Tamarin
        to formally verify secrecy and authentication properties.
  \item \textbf{WebSocket transport}: Replacing HTTP polling with
        a persistent WebSocket connection for lower latency and
        server-push transfer updates.
  \item \textbf{End-to-end encryption}: Per-file client-side
        encryption keys to achieve zero-knowledge server storage.
  \item \textbf{Algorithm agility}: Handshake negotiation supporting
        multiple PQC KEM/signature schemes (NTRU, McEliece) to future-proof
        the protocol against individual algorithm weaknesses.
\end{itemize}
