\chapter{Contribution}
\lhead{Chapter 5. \emph{Contribution}}
\label{chap.5}

This chapter presents the key contributions of the Quantum Secure File Transfer Protocol (Q-SFTP) project, highlighting the novel aspects and significant technical advancements achieved in the implementation.

\section{Innovations of the Proposed Method}

The novelty of this project lies in the design and implementation of a complete, production-grade post-quantum file transfer system. While prior works have focused on individual cryptographic algorithms or theoretical protocol designs, Q-SFTP uniquely combines the following innovations:

\begin{enumerate}
    \item \textbf{Full PQC Integration:} Complete replacement of classical cryptographic primitives with NIST-standardized algorithms:
    \begin{itemize}
        \item \textbf{CRYSTALS-Kyber512} (FIPS 203 / ML-KEM) for key encapsulation, providing 128-bit post-quantum security.
        \item \textbf{CRYSTALS-Dilithium2} (FIPS 204 / ML-DSA) for digital signatures, providing tamper-proof mutual authentication.
        \item \textbf{AES-256-GCM} for symmetric encryption with authenticated integrity.
    \end{itemize}
    
    \item \textbf{Four-Phase Mutual Authentication:} A novel handshake protocol achieving mutual authentication through certificate exchange, Kyber key encapsulation, transcript hashing, and dual Dilithium signatures — all within four round-trips.
    
    \item \textbf{BLAKE2b-256 Integrity Ecosystem:} Unlike simple checksum approaches, Q-SFTP implements a comprehensive integrity ecosystem with dual-side hash verification (client and server), a persistent hash registry, and a background integrity checker that periodically re-verifies all stored files.
    
    \item \textbf{Chunked Resumable Transfers over PQC Channels:} A chunked transfer protocol that operates within the PQC-encrypted channel, with persistent state tracking on both client (localStorage) and server (JSON log), enabling pause/resume across sessions and connection failures.
    
    \item \textbf{Privacy-Preserving File Storage:} Automatic metadata stripping for images (EXIF), PDFs, and DOCX files before storage, combined with IP anonymization in activity logs using daily rotating salts.
    
    \item \textbf{Lightweight Post-Quantum CA:} A self-contained Certificate Authority built entirely on Dilithium2, without dependency on classical X.509 infrastructure.
\end{enumerate}

\section{Key Technical Contributions}

\subsection{Modular Security Architecture}

The four-tier modular architecture (Client $\rightarrow$ Flask $\rightarrow$ Handshake Server $\rightarrow$ Storage) separates concerns such that cryptographic primitives can be upgraded independently of the application logic. This design anticipates future NIST standard revisions or the adoption of alternative PQC algorithms.

\subsection{Defense-in-Depth File Protection}

Q-SFTP implements multiple layers of file protection:

\begin{table}[H]
    \centering
    \caption{Defense-in-depth layers for file protection}
    \label{tab:defense}
    \begin{tabular}{|l|l|p{7cm}|}
        \hline
        \textbf{Layer} & \textbf{Module} & \textbf{Protection} \\
        \hline
        Transport & Handshake Server & AES-256-GCM encryption with Kyber-derived session keys \\
        \hline
        Validation & FileValidator & Extension blacklisting (17 types) + magic number verification (20+ formats) \\
        \hline
        Integrity & HashVerifier & Dual-side BLAKE2b-256 + persistent registry + periodic background checks \\
        \hline
        Privacy & PrivacyManager & EXIF/PDF/DOCX metadata stripping + IP anonymization \\
        \hline
        Access & AuthManager + RBAC & Three-tier role enforcement + directory isolation \\
        \hline
        Audit & ActivityLogger & Tamper-evident log chain with hash chaining \\
        \hline
    \end{tabular}
\end{table}

\subsection{Production-Ready Web Interface}

Unlike most PQC research prototypes that provide only CLI interfaces, Q-SFTP includes a fully functional web GUI that:
\begin{itemize}
    \item Abstracts all cryptographic complexity behind an intuitive FTP-style interface.
    \item Provides real-time progress bars for chunked transfers with pause/resume controls.
    \item Includes an admin panel for user management, activity log viewing, and integrity monitoring.
    \item Supports dark/light theme switching and responsive design.
\end{itemize}

\section{Novel Findings}

During development and testing, several noteworthy findings emerged:

\begin{itemize}
    \item \textbf{Certificate Size Optimization:} Post-quantum certificates are inherently larger than classical ones (Dilithium2 public keys are 1,312 bytes vs. RSA-2048's 256 bytes). However, the one-time cost during handshake does not significantly impact overall transfer performance.
    
    \item \textbf{BLAKE2b Performance Advantage:} Client-side BLAKE2b computation using the \texttt{blakejs} library proved approximately 3$\times$ faster than SHA-256 via the WebCrypto API for file hashing, confirming its suitability for real-time integrity verification.
    
    \item \textbf{Resumable Transfer Resilience:} Testing demonstrated that chunked transfers with 256 KB chunks and localStorage persistence can reliably survive connection interruptions, browser refreshes, and even system restarts, with minimal overhead ($\approx 0.01\%$ for cryptographic framing).
    
    \item \textbf{Protocol Adaptability:} The modular architecture allowed the complete replacement of the original DH-based key exchange with Kyber512 without modifying any file management or UI code — validating the separation-of-concerns design.
\end{itemize}

\section{Comparison with Existing Solutions}

\begin{table}[H]
    \centering
    \caption{Feature comparison: Q-SFTP vs. existing approaches}
    \label{tab:comparison}
    \begin{tabular}{|l|c|c|c|}
        \hline
        \textbf{Feature} & \textbf{Classical SFTP} & \textbf{Hybrid PQ-TLS} & \textbf{Q-SFTP (Ours)} \\
        \hline
        Quantum-safe KEM & $\times$ & \checkmark & \checkmark \\
        \hline
        Quantum-safe Signatures & $\times$ & Partial & \checkmark \\
        \hline
        Mutual Authentication & \checkmark & \checkmark & \checkmark \\
        \hline
        File Integrity (BLAKE2b) & $\times$ & $\times$ & \checkmark \\
        \hline
        Resumable Transfers & $\times$ & $\times$ & \checkmark \\
        \hline
        Metadata Stripping & $\times$ & $\times$ & \checkmark \\
        \hline
        RBAC + Isolation & Partial & Partial & \checkmark \\
        \hline
        Web GUI & $\times$ & $\times$ & \checkmark \\
        \hline
        Admin Panel & $\times$ & $\times$ & \checkmark \\
        \hline
    \end{tabular}
\end{table}

In total, the contributions of this work go beyond cryptographic algorithm integration to deliver a comprehensive, deployable system with security, usability, and reliability features necessary for practical adoption.
